Inferring the effective start dates of non-pharmaceutical interventions during COVID-19 outbreaks

### 2020-05-20 ms/ms.tex
typos in ms.tex: 12 regions; added digits to csv reports

%%% Model selection %%%
\paragraph*{Model comparison.}
We perform model selection using two methods.
First, we compute WAIC (widely applicable information criterion)~\citep{gelman2013bayesian},
\begin{equation} \label{eq:WAIC}
\begin{aligned}
\mathit{WAIC}(\theta, \vec{X}) &= -2\log\mathbb{E}[\mathcal{L}(\theta \mid \vec{X})] + 2\mathbb{V}[\log\mathcal{L}(\theta \mid \vec{X})]
\end{aligned}
\end{equation}
where $\mathbb{E}[\cdot]$ and $\mathbb{V}[\cdot]$ are the expectation and variance operators taken over the posterior distribution $P(\theta \mid \vec{X})$.
We compare models by reporting their relative WAIC; lower is better~(\autoref{table:WAIC}).
A minority of MCMC chains that fail to fully converge can lead to overestimation of the variance (the second term in \autoref{eq:WAIC}).
Therefore, we exclude from the WAIC computation chains with mean log-likelihood that is three standard deviations or more from the overall mean.

@@

%%% Fig 1 %%%
\begin{figure}[h]
    \centering
	\includegraphics[width=0.6\textwidth]{../figures/Fig1.pdf}
    \caption{
    \textbf{Official and effective start of non-pharmaceutical interventions.}
    	The difference between $\hat{\tau}$ the effective and $\tau^*$ the official start of NPI is shown for different regions. The effective NPI dates in Italy and Wuhan are significantly delayed compared to the official dates, whereas in Denmark, France, Spain, and Germany, the effective date is earlier than the official date.
	$\hat{\tau}$ is the posterior median, see \autoref{table:estimated-params}. $\tau^*$ is the last NPI date, see \autoref{table:NPI_dates}. Thin and bold lines show 95\% and 75\% credible intervals, respectively (area in which $P(|\tau - \hat{\tau}| \mid \vec{X}) = 0.95$ and $0.75$.)
    }
    \label{fig:fig1}
\end{figure}

@@

We compare the official $\tau^*$ and effective $\hat{\tau}$ start of NPIs and find that in most regions the effective start of NPI  significantly differs from the official date (\autoref{fig:fig1}),; that is, the credible interval on $\hat{\tau}$ does not include $\tau^*$ (\autoref{fig:fig1}). The exceptions are, as with the comparison to the simpler models, Austria, Belgium, and United Kingdom, as well another country (see below). Norway also has a relatively wide credible interval, which could be expected as it has the longest duration between the first and last NPIs~(\autoref{table:NPI_dates}).
In the following, we describe our findings on late and early effective start of NPI in detail.

@@

In Italy, the first case was officially confirmed on Feb~21. School closures were implemented on Mar~5~\citep{Flaxman2020}, a lockdown was declared in Northern Italy on Mar~8, with social distancing implemented in the rest of the country, and the lockdown was extended to the entire nation on Mar 11~\citep{Gatto2020}.
That is, the first and last official dates are Mar~8 and Mar~11.
However, we estimate the effective date $\hat{\tau}$ at Mar~16 ($\pm$0.47 days 95\% CI ; \autoref{fig:late}).

Similarly, in Wuhan, China, a lockdown was ordered on Jan 23~\citep{Li2020}, but we estimate the effective start of NPIs to be several days layer at Feb~2 ($\pm$2.85 days 95\% CI \autoref{fig:late}).

@@

%% Table 2 %%%
\begin{landscape}
\begin{table}[h]
\centering
\input{../figures/Table-estimated-params.tex}
\caption{
\textbf{Parameter estimates for different regions.}
See \autoref{eq:model} for model parameters.
All estimates are posterior medians.
75\% and 95\% credible intervals given for $\tau$, in days.
$\tau^*$ is the official last NPI date, see \autoref{table:NPI_dates}.
}
\label{table:estimated-params}
\end{table}
\end{landscape}

>>>>

%%% Model selection %%%
\paragraph*{Model comparison.}
We perform model selection using two methods.
First, we compute WAIC (widely applicable information criterion)~\citep{gelman2013bayesian},
\begin{equation} \label{eq:WAIC}
\begin{aligned}
WAIC(\theta, \vec{X}) &= -2\log\mathbb{E}[\mathcal{L}(\theta \mid \vec{X})] + 2\mathbb{V}[\log\mathcal{L}(\theta \mid \vec{X})]
\end{aligned}
\end{equation}
where $\mathbb{E}[\cdot]$ and $\mathbb{V}[\cdot]$ are the expectation and variance operators taken over the posterior distribution $P(\theta \mid \vec{X})$.
We compare models by reporting their relative WAIC; lower is better~(\autoref{table:WAIC}).
A minority of MCMC chains that fail to fully converge can lead to overestimation of the variance (the second term in \autoref{eq:WAIC}).
Therefore, we exclude from the WAIC computation chains with mean log-likelihood that is three standard deviations or more from the overall mean.

@@

%%% Fig 1 %%%
\begin{figure}[h]
    \centering
	\includegraphics[width=0.6\textwidth]{../figures/Fig1.pdf}
    \caption{
    \textbf{Official and effective start of non-pharmaceutical interventions.}
    	The difference between $\hat{\tau}$ the effective and $\tau^*$ the official start of NPI is shown for different regions. The effective NPI dates in Italy and Wuhan are significantly delayed compared to the official dates, whereas in Denmark, France, Spain, and Germany, the effective date is earlier than the official date.
	$\hat{\tau}$ is the posterior median, see \autoref{table:estimated-params}. $\tau^*$ is the last NPI date, see \autoref{table:NPI_dates}. Thin and bold lines show 95\% and 75\% credible intervals (area in which $P(|\tau - \hat{\tau}| \mid \vec{X}) = 0.95$ and $0.75$.)
    }
    \label{fig:fig1}
\end{figure}

@@

We compare the official $\tau^*$ and effective $\hat{\tau}$ start of NPIs and find that in most regions the effective start of NPI  significantly differs from the official date (\autoref{fig:fig1}), that is, the credible interval on $\hat{\tau}$ does not include $\tau^*$ (\autoref{fig:fig1}). The exceptions are, as with the comparison to the simpler models, Austria, Belgium, and United Kingdom, as well another country (see below). Norway also has a relatively wide credible interval, which could be expected as it has the longest duration between the first and last NPIs~(\autoref{table:NPI_dates}).
In the following, we describe our findings on late and early effective start of NPI in detail.

@@

In Italy, the first case was officially confirmed on Feb~21. School closures were implemented on Mar~5~\citep{Flaxman2020}, a lockdown was declared in Northern Italy on Mar~8, with social distancing implemented in the rest of the country, and the lockdown was extended to the entire nation on Mar 11~\citep{Gatto2020}.
That is, the first and last official dates are Mar~8 and Mar~11.
However, we estimate the effective date $\hat{\tau}$ at Mar~16 ($\pm$0.47 days 95\% CI ; \autoref{fig:late}).
Similarly, in Wuhan, China, a lockdown was ordered on Jan 23~\citep{Li2020}, but we estimate the effective start of NPIs to be several days layer at Feb~2 ($\pm$2.85 days 95\% CI \autoref{fig:late}).

@@

%% Table 2 %%%
\begin{landscape}
\begin{table}[h]
\centering
\input{../figures/Table-estimated-params.tex}
\caption{
\textbf{Parameter estimates for different regions.}
See \autoref{eq:model} for model parameters.
All estimates are posterior medians.
75\% and 95\% credible intervals given only for $\tau$, in days.
$\tau^*$ is the official last NPI date, see \autoref{table:NPI_dates}.
}
\label{table:estimated-params}
\end{table}
\end{landscape}
